---
title: "Wrangle USGS"
author: "John Fay"
date: "Spring 2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Objectives
- Compare water supply and usage rates (mgal/day) to demographic metrics (e.g. SVI)
  - Compute rates for a given year (or yearly avgs) for each county
    | FIPS | year | Metric | MGal per Day |

- Compare water supply and usage changes over time to demographic metrics

```{r Setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dataRetrieval)
```


#### 1. Read in the NWIS water usage data [later: via the API]
```{r retrieve.usage.data}
#From the dataRetrieval API
df_use <- readNWISuse('NC','ALL',c(2000,2005,2010,2015),'ALL') %>%  # Read data for NC
  mutate(FIPS = paste0(state_cd,county_cd), .before=1)  %>%         # Add FIPS column
  select(-c(state_cd,state_name,county_cd,county_nm)) %>%           # Remove labels
  replace(.=='-',0) %>%                                             # Convert missing values to zero
  mutate_if(is.character,as.numeric) %>%                            # Convert data fields to numeric
  mutate(year = as.factor(year),FIPS=as.factor(FIPS))               # Revert year & FIPS to factor

```         
   

#### 2. Translate raw columns to accounting table categories
```{r} 
df_year <- df_use %>% 
  group_by(year,FIPS) %>% 
  summarize(#Public Water Supply withdrawals
            WaterSupply.Surface.Fresh =  sum(Public.Supply.self.supplied.surface.water.withdrawals..fresh..in.Mgal.d),  #B7
            WaterSupply.Surface.Saline = sum(Public.Supply.self.supplied.surface.water.withdrawals..saline..in.Mgal.d), #B8
            WaterSupply.Surface.Total =  sum(Public.Supply.self.supplied.surface.water.withdrawals..fresh..in.Mgal.d,   #B5
                                             Public.Supply.self.supplied.surface.water.withdrawals..saline..in.Mgal.d),
            WaterSupply.Ground.Fresh =   sum(Public.Supply.self.supplied.groundwater.withdrawals..fresh..in.Mgal.d),    #B10
            WaterSupply.Ground.Saline =  sum(Public.Supply.self.supplied.groundwater.withdrawals..saline..in.Mgal.d),   #B11
            WaterSupply.Ground.Total =   sum(Public.Supply.self.supplied.groundwater.withdrawals..fresh..in.Mgal.d,     #B9
                                             Public.Supply.self.supplied.groundwater.withdrawals..saline..in.Mgal.d),
            WaterSupply.Total.Abstraction = sum(WaterSupply.Surface.Total, WaterSupply.Ground.Total),                   #B5
            WaterSupply.Total.TotalUse = sum(WaterSupply.Surface.Total, WaterSupply.Ground.Total),                      #B16
            WaterSupply.OtherUnits = 0, 
            
            #Sewage Treatment withdrawals (NOT USED)
            
            #Household withdrawals (wells)
            Household.Surface.Fresh     = sum(Domestic.self.supplied.surface.water.withdrawals..fresh..in.Mgal.d),      #D7
            Household.Surface.Saline    = sum(Domestic.self.supplied.surface.water.withdrawals..saline..in.Mgal.d),     #D8
            Household.Ground.Fresh      = sum(Domestic.self.supplied.groundwater.withdrawals..fresh..in.Mgal.d),        #D10
            Household.Ground.Saline     = sum(Domestic.self.supplied.groundwater.withdrawals..saline..in.Mgal.d),       #D11
            Household.Total.Abstraction = sum(Household.Surface.Fresh,Household.Surface.Saline,                         #D5
                                              Household.Ground.Fresh, Household.Ground.Saline),
            Household.OtherUnts = sum(Domestic.total.self.supplied.withdrawals.plus.deliveries..in.Mgal.d -             #D13
                                      Domestic.self.supplied.surface.water.withdrawals..fresh..in.Mgal.d - 
                                      Domestic.self.supplied.surface.water.withdrawals..saline..in.Mgal.d -
                                      Domestic.self.supplied.groundwater.withdrawals..fresh..in.Mgal.d - 
                                      Domestic.self.supplied.groundwater.withdrawals..saline..in.Mgal.d),
            Household.TotalUse = sum(Household.OtherUnts, Household.Total.Abstraction),                                 #D16
            
            #Irrigation
            Irrigation.Crop.Total.Abstraction = sum(                                                                    #E5
              Irrigation..Crop.self.supplied.surface.water.withdrawals.for.crops..fresh..in.Mgal.d,
              Irrigation..Crop.self.supplied.groundwater.withdrawals.for.crops..fresh..in.Mgal.d),
            Irrigation.Crop.TotalUse = Irrigation.Crop.Total.Abstraction,                                               #E16
            Irrigation.Golf.Total.Abstraction = sum(
              Irrigation..Golf.Courses.self.supplied.surface.water.withdrawals.for.golf.courses..fresh..in.Mgal.d,
              Irrigation..Golf.Courses.self.supplied.groundwater.withdrawals.for.golf.courses..fresh..in.Mgal.d),
            Irrigation.Golf.TotalUse = Irrigation.Golf.Total.Abstraction,
            
            #Animal production
            Animal.Total.Abstraction = sum(
              Livestock.self.supplied.surface.water.withdrawals..fresh..in.Mgal.d,
              Livestock.self.supplied.groundwater.withdrawals..fresh..in.Mgal.d),
            Animal.TotalUse = Animal.Total.Abstraction,
            
            #Aquaculture
            Aquaculture.Total.Abstraction = sum(
              Aquaculture.total.self.supplied.withdrawals..surface.water..in.Mgal.d,
              Aquaculture.total.self.supplied.withdrawals..groundwater..in.Mgal.d
            ),
            Aquaculture.TotalUse = Aquaculture.Total.Abstraction,
            
            #Manufacturing
            Manufacturing.Total.Abstraction = sum(
              Industrial.total.self.supplied.withdrawals..surface.water..in.Mgal.d,
              Industrial.total.self.supplied.withdrawals..groundwater..in.Mgal.d
            ),
            Manufacturing.TotalUse = Manufacturing.Total.Abstraction,
            
            #Mining
            Mining.Total.Abstraction = sum(
              Mining.total.self.supplied.withdrawals..surface.water..in.Mgal.d,
              Mining.total.self.supplied.withdrawals..groundwater..in.Mgal.d
            ),
            Mining.TotalUse =  Mining.Total.Abstraction
            
            #Power
            
            #End summarize
            ) 

```


#### 3. Read in the CDC SVI data
```{r read.in.SVI.data}
#Fetch in the raw data
df_svi <- read_csv('https://svi.cdc.gov/Documents/Data/2018_SVI_Data/CSV/SVI2018_US_COUNTY.csv') %>% 
  filter(ST == '37') %>% 
  select(FIPS, starts_with("E_"))%>% 
  mutate(FIPS = as.factor(FIPS))

#Fetch the data dictionary
data_dict <- read_csv("https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/csv/SVIDocumentation_Table_DataDictionary_2018.csv") %>% 
  select(`2018 VARIABLE NAME`, `2018 DESCRIPTION`) %>% 
  filter(str_detect(`2018 VARIABLE NAME`, "\\bE")) 

#Create lists of the SVI fields and their new name
svi_fields <- colnames(select(df_svi,starts_with("E_"))) #colnames(df_svi)[-1]

#Function to pull new name from old name in data dictionary
update_name <- function(colName){
  new_name <- data_dict %>% 
    filter(`2018 VARIABLE NAME` == !!colName) %>% 
    select(`2018 DESCRIPTION`)
  new_name <- new_name[[1]]
  return (new_name)}

#Look up new names
alt_names <- svi_fields %>% #data_dict$`2018 VARIABLE NAME` %>% 
  map(update_name)

#Rename fields in df_svi
df_svi <- setnames(df_svi, old=svi_fields, new = unlist(alt_names))

```


```{r join.and.save}
#Join the usage and svi tables on FIPS
df_all <- left_join(df_year,df_svi,by=NULL)

write_csv(df_all,'water_svi_data.csv')
```

